Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2017-06-06 paper/submitted/round2_png/algorithms_v4_cmame.tex
more edits...

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Error Estimate Localization}\label{sec:errLocal}
%------------------------------------------------------------------------------------------------------------------------%
\Cref{alg:refSeries} does not require a specific method for localizing the error estimate. A na\"{i}ve approach would be to write the error estimate as a sum of integrals over elements and their boundaries, and calculate the error contribution by each element as the integral over that element. While simple, this method can lead to non-zero error contributions from elements in which the high-fidelity model is already being used, making the error decomposition more difficult to interpret and use for refinement.

We instead use the alternative method described in \cite{vanOpstaletal15}, decomposing the error estimate into contributions from locally supported basis functions rather than elements. For ease of implementation, we can choose these basis functions to be the same as those spanning the space in which we solve for the primary, auxiliary, and supplementary adjoint variables. Near the interfaces between the low-fidelity and high-fidelity regions, basis functions may have their support divided between the two regions and thus have a nonzero error contribution. The basis functions with the largest error contributions are flagged, and the elements in their support are added to the high-fidelity subdomain. Basis functions whose support lie completely in the high-fidelity region will have a zero contribution and the elements in their support are never flagged for refinement again.

>>>>

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Error Estimate Localization}\label{sec:errLocal}
%------------------------------------------------------------------------------------------------------------------------%

\Cref{alg:refSeries} does not require a specific method for localizing the error estimate. A na\"{i}ve approach would be to write the error estimate as a sum of integrals over elements and their boundaries, and calculate the error contribution by each element as the integral over that element. While simple, this method can lead to non-zero error contributions from elements in which the high-fidelity model is already being used, making the error decomposition more difficult to interpret and use for refinement.

We instead use the alternative method described in \cite{vanOpstaletal15}, decomposing the error estimate into contributions from locally supported basis functions rather than elements. Recall the error estimate 
%
\begin{equation}
\epsilon=-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{MF})(\Lambda)+\mathcal M_{HF}(\Psi_{MF})-\mathcal M_{MF}(\Psi_{MF})
\end{equation}
%
from \Cref{alg:refSeries}. Let $(Q^h\times U^h\times U^h)^3 \subset (Q\times U\times U)^3$ be the finite-dimensional conforming subspace in which we solve for the approximations $\Lambda^h$ and $\Psi_{MF}^h$; we seek to decompose the error estimate
%
\begin{equation}
\epsilon^h=-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{MF}^h)(\Lambda^h)+\mathcal M_{HF}(\Psi_{MF}^h)-\mathcal M_{MF}(\Psi_{MF}^h).
\end{equation}
%
Define a basis $\Phi^h=\{(\varphi,\upvarphi)_i\}_{i\in I}$ consisting of locally supported functions such that $\textrm{span}(\Phi^h)=(Q^h\times U^h\times U^h)^3$, so that we can write $(\Lambda^h,\Psi_{MF}^h)=(\sum\limits_{i\in I}\varphi_i\lambda_i,\sum\limits_{i\in I}\upvarphi_i \psi_i)$. Then the error estimate satisfies
%
\begin{equation}
\epsilon^h \leq \sum_{i\in I} \varepsilon^h_i,
\end{equation}
%
where,
%
\begin{equation}\label{eq:basisblame}
\varepsilon^h_i = \left| -\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{MF})(\lambda_i\varphi_i)+\mathcal M_{HF}(\upvarphi_i \psi_i)-\mathcal M_{MF}(\upvarphi_i \psi_i) \right|
\end{equation}
%
can be interpreted as the error contribution from the basis function $(\varphi,\upvarphi)_i$. Near the interfaces between the low-fidelity and high-fidelity regions, basis functions may have their support divided between the two regions and thus have a nonzero error contribution. The basis functions with the largest error contributions are flagged, and the elements in their support are added to the high-fidelity subdomain. Basis functions whose support lie completely in the high-fidelity region will have a zero contribution and the elements in their support are never flagged for refinement again.

### 2017-06-06 paper/submitted/round2_png/numerical_experiments_v4_cmame.tex
more edits...

For the numerical simulations, we use the finite element method (FEM), employing a continuous Galerkin formulation with linear Lagrange elements. We use the \texttt{libMesh} library~\cite{libMeshPaper} for the FEM calculations.
%\red{The library offers easy calculation of adjoint systems, error estimates and subdomain restricted variables.}\footnote{But we don't use Libmesh's error estimates or restrict variables to subdomains...}
The domain is discretized by a regular mesh of quadrilaterals, with 250 and 50 elements along the $x_1$ and $x_2$ directions, respectively, for a total of 12,500 elements, resulting in 12,801 degrees of freedom per variable; all variables are represented by the same basis. The diffusion coefficient is such that the cell P\'{e}clet number never exceeds 0.1, and thus no stabilization is required.

@@

We consider the same high-fidelity model as in \Cref{sec:cdvcdr}:
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in U,
\end{equation}
with the same diffusion coefficient $k_d = 0.1$  and reaction coefficient $k_r = -42$. The low-fidelity model
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in\R
\end{equation}
differs from the high-fidelity model only in that the parameter $q$ is a constant instead of a field; $f(q)$ is a constant function equal to $q$ everywhere in the domain. The intermediate mixed-fidelity models thus have parameter fields (and forcing functions) that are constant over the subregions of the domain where the low-fidelity model is used; disconnected low-fidelity regions still share the same parameter value. For ease of implementation, we require that the resulting parameter field remain continuous at the interface between the low-fidelity and high-fidelity subdomains, although this constraint is not necessary for the theory to hold. The domain, mesh, boundary conditions, and velocity field, as well as the observations and QoI, remain the same as described in \Cref{sec:cdvcdr}; we continue to infer for the forcing field. The forcing field (and its auxiliary and supplementary adjoint conterparts) are represented with the same basis functions as the other variables in the high-fidelity regions, and forced to be constant in the low-fidelity regions. As the inverse problem is ill-posed, except when the low-fidelity model is used throughout the domain, regularization is used; the Tikhonov regularization term in \Cref{eq:invOpt_obj} is $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2+f(q)^2\:\textrm{d}A$, where $\beta=10^{-3}$ is a regularization coefficient.

@@

The corresponding true and estimated absolute errors in the QoI are shown in \Cref{fig:svfErr}.
%
\begin{figure}[htbp]
\centering
%\includegraphics[width=0.8\textwidth]{svf/err_est.pdf}
\includegraphics[width=0.8\textwidth]{svf/err_est_v2.pdf}
\caption{True and estimated absolute relative error in QoI, plotted as a function of the percentage area of the domain in which the high-fidelity field-parameter model is used. The dotted line indicates the target relative QoI error.}
\label{fig:svfErr}
\end{figure}
%
In this case, we see that we must use the high-fidelity model in most of the domain in order to get an accurate QoI. The adaptive algorithm requires us to use the field representation of the high-fidelity model in much of the left half of the domain; this reflects the topology of the inferred parameter field in the high-fidelity inverse problem, which is only relatively constant towards the right portion of the domain. We also see that in this case, in contrast to the example in \Cref{sec:cdvcdrBaseRef}, increasing the proportion of the domain in which the high-fidelity model is used does not monotonically decrease the error in the QoI.

@@

\Cref{tab:ref3D_diffmesh} shows the error at the end of each adaptive iteration. Each iteration of the adaptive algorithm uses the solution of the previous iteration as its initial guess. The number of degrees of freedom of each of the primary (and, if applicable, auxiliary) variables at each iteration is also given; the supplementary adjoint is solved on the high-fidelity mesh and thus each of its variables has the same number of degrees of freedom as each of the primary variables in the high-fidelity inverse problem. The high-fidelity inverse problem does not converge when the low-fidelity solution is used as an initial guess. Instead, we solve the inverse problem for an intermediate model
%
\begin{equation}
\nabla\cdot(n\vec{V}u - nD\nabla u) + k_ru^2 = f(q)
\end{equation}
%
and use the solution as an initial guess for the high-fidelity problem; equivalently, we use two steps of natural continuation on the reaction parameter: $k_r=0$, then $k_r=4.2\cdot10^{-4}$.
%
\begin{table}[htbp]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
\multirow{2}{*}{Case} & \multirow{2}{*}{$\%$HF} & \multirow{2}{*}{DOFs} & \multirow{2}{*}{QoI} & Error & Error & $\%$ Relative \\
& & & & (Estimated) & (Actual) & Error (Actual)  \\ \hline
LF   & 0    & 9537  & 168710 & -16463 & -85663 & -103    \\
MF01 & 5.2  & 13417 & 167366 & -7207  & -84319 & -102    \\
MF02 & 11.4 & 17895 & 89777  & -6208  & -6730  & -8.10   \\
MF03 & 16.3 & 21001 & 85880  & -2473  & -2833  & -3.41   \\
MF04 & 22.0 & 24528 & 83902  & -711   & -855   & -1.03   \\
MF05 & 27.7 & 27984 & 83119  & 32     & -72    & -0.087  \\
HF   & 100  & 70785 & 83047  & --     & --     & --    \\ \hline
\end{tabular}
\caption{Runtime and relative errors of adaptive algorithm iterations given relative error tolerance of $0.1\%$; relative errors are given with respect to the high-fidelity QoI estimate.}
\label{tab:ref3D_diffmesh}
\end{table}
%

>>>>

For the numerical simulations, we use the finite element method (FEM), employing a continuous Galerkin formulation with first-order Lagrange elements. We use the \texttt{libMesh} library~\cite{libMeshPaper} for the FEM calculations.
%\red{The library offers easy calculation of adjoint systems, error estimates and subdomain restricted variables.}\footnote{But we don't use Libmesh's error estimates or restrict variables to subdomains...}
The domain is discretized by a regular mesh of quadrilaterals, with 250 and 50 elements along the $x_1$ and $x_2$ directions, respectively, for a total of 12,500 elements, resulting in 12,801 degrees of freedom per variable; all variables are represented by the same basis. The diffusion coefficient is such that the cell P\'{e}clet number never exceeds 0.1, and thus no stabilization is required.

@@

We consider the same high-fidelity model as in \Cref{sec:cdvcdr}:
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in H^1(\Omega),
\end{equation}
with the same diffusion coefficient $k_d = 0.1$  and reaction coefficient $k_r = -42$. The low-fidelity model
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2= f(q),\quad q\in U_{LF}
\end{equation}
differs from the high-fidelity model only in that $U_{LF}$ is limited to the space of constant functions over $\Omega$; we maintain $f(q)=q$. The intermediate mixed-fidelity models thus have parameter fields (and forcing functions) that are constant over the subregions of the domain where the low-fidelity model is used; disconnected low-fidelity regions still share the same parameter value. The domain, mesh, boundary conditions, and velocity field, as well as the observations and QoI, remain the same as described in \Cref{sec:cdvcdr}; we continue to infer for the forcing field. For ease of implementation, we require that the resulting parameter field remain continuous at the interface between the low-fidelity and high-fidelity subdomains, although this constraint is not necessary for the theory to hold. Also for ease of implementation and for localizing the error, the forcing field and its auxiliary and supplementary adjoint conterparts are represented with the same basis functions as the other variables in the high-fidelity regions, and forced to be constant in the low-fidelity regions using a penalty term. Although the chosen pair of models has the potential to give mixed-fidelity models with fewer degreees of freedom in the parameter, this is not the case in our implementation; however, we are not looking to demonstrate computational savings with this example, but rather to explore the behavior of \Cref{alg:refSeries} when applied to such a pair of models. As the inverse problem is ill-posed, except when the low-fidelity model is used throughout the domain, regularization is used; the Tikhonov regularization term in \Cref{eq:invOpt_obj} is $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2+f(q)^2\:\textrm{d}A$, where $\beta=10^{-3}$ is a regularization coefficient.

@@

\Cref{fig:svfRefParams} shows the inferred parameters using some of the mixed-fidelity models and the high-fidelity model.
%
\begin{figure}[htbp]
\centering
\subfloat[MF$_1$ ($10\%$ HF)]{
	\includegraphics[width=0.8\textwidth]{svf/svf_param_MF01_clim_HF_whitebg.png}
	\label{subfig:svfparamMF1}
} \\
\subfloat[MF$_3$ ($30\%$ HF)]{
	\includegraphics[width=0.8\textwidth]{svf/svf_param_MF03_clim_HF_whitebg.png}
	\label{subfig:svfparamMF3}
} \\
\subfloat[MF$_6$ ($60\%$ HF)]{
  \includegraphics[width=0.8\textwidth]{svf/svf_param_MF06_clim_HF_whitebg.png}
  \label{subfig:svfparamMF6}
} \\
\subfloat[HF]{
	\includegraphics[width=0.8\textwidth]{svf/svf_params_HF_whitebg.png}
	\label{subfig:svfparamHF}
}
\caption{Inferred parameters using mixed-fidelity (\protect\subref{subfig:svfparamMF1}--\protect\subref{subfig:svfparamMF6}) and high-fidelity (\protect\subref{subfig:svfparamHF}) models. All figures share the color-scale for the high-fidelity plot.}
\label{fig:svfRefParams}
\end{figure}
%
The corresponding true and estimated absolute errors in the QoI are shown in \Cref{fig:svfErr}.
%
\begin{figure}[htbp]
\centering
%\includegraphics[width=0.8\textwidth]{svf/err_est.pdf}
\includegraphics[width=0.8\textwidth]{svf/err_est_v2.pdf}
\caption{True and estimated absolute relative error in QoI, plotted as a function of the percentage area of the domain in which the high-fidelity field-parameter model is used. The dotted line indicates the target relative QoI error.}
\label{fig:svfErr}
\end{figure}
%
In this case, we see that we must use the high-fidelity model in most of the domain in order to get an accurate QoI. The adaptive algorithm requires us to use the field representation of the high-fidelity model in much of the left half of the domain; this reflects the topology of the inferred parameter field in the high-fidelity inverse problem, which is only relatively constant towards the right portion of the domain (compare \Cref{subfig:svfparamMF6} and \Cref{subfig:svfparamHF}). We also see that in this case, in contrast to the example in \Cref{sec:cdvcdrBaseRef}, increasing the proportion of the domain in which the high-fidelity model is used does not monotonically decrease the error in the QoI. Note that although the inferred parameters in the first and third mixed-fidelity models (see \Cref{subfig:svfparamMF1} and \Cref{subfig:svfparamMF3}) are very different, they give very similar QoI error; the same is true for the second and fourth mixed-fidelity models. These provide additional examples supporting the idea that using inferring parameters with models of different mixtures of fidelities can produce similar QoIs.

%compare MF01 and MF03, MF06 and HF
%1 and 3 look very different but have very similar true errors - different params can give similar qoi, though in this case the qoi is not close to HF; more HF doesn't necessarily decrease error, interplay between having freedom to match data and all LF regions being flat...
%6 and HF look very similar; most of HF is very flat

@@

\Cref{tab:ref3D_diffmesh} shows the error at the end of each adaptive iteration. Each iteration of the adaptive algorithm uses the solution of the previous iteration as its initial guess. The number of degrees of freedom of each of the primary (and, if applicable, auxiliary) variables at each iteration is also given; the supplementary adjoint is solved on the high-fidelity mesh and thus each of its variables has the same number of degrees of freedom as each of the primary variables in the high-fidelity inverse problem. The high-fidelity inverse problem does not converge when the low-fidelity solution is used as an initial guess. Instead, we solve the inverse problem for an intermediate model
%
\begin{equation}
\nabla\cdot(n\vec{V}u - nD\nabla u) + k_ru^2 = f(q)
\end{equation}
%
with $k_r=0$, and use the solution as an initial guess for the high-fidelity problem; this is equivalent to using two steps of natural continuation on the reaction parameter: $k_r=0$, then $k_r=4.2\cdot10^{-4}$.
%
\begin{table}[htbp]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
\multirow{2}{*}{Case} & \multirow{2}{*}{$\%$HF} & \multirow{2}{*}{DOFs} & \multirow{2}{*}{QoI} & Error & Error & $\%$ Relative \\
& & & & (Estimated) & (Actual) & Error (Actual)  \\ \hline
LF   & 0    & 9537  & 168710 & -16463 & -85663 & -103    \\
MF01 & 5.2  & 13417 & 167366 & -7207  & -84319 & -102    \\
MF02 & 11.4 & 17895 & 89777  & -6208  & -6730  & -8.10   \\
MF03 & 16.3 & 21001 & 85880  & -2473  & -2833  & -3.41   \\
MF04 & 22.0 & 24528 & 83902  & -711   & -855   & -1.03   \\
MF05 & 27.7 & 27984 & 83119  & 32     & -72    & -0.087  \\
HF   & 100  & 70785 & 83047  & --     & --     & --    \\ \hline
\end{tabular}
\caption{Runtime and relative errors of adaptive algorithm iterations given relative error tolerance of $0.1\%$; relative errors are given with respect to the high-fidelity QoI estimate.}
\label{tab:ref3D_diffmesh}
\end{table}
%
